New orbits of irregular satellites based on thousands of new observations and predictions for stellar occultations up to 2020

### 2016-05-31 Irregular-occ/Irregular-occ.tex
.

\title[New orbits and predictions of stellar occultations by irregular satellites up to 2020]{New orbits for irregular satellites based on thousands of new observations and predictions of stellar occultations up to 2020}
\author[A. R. Gomes-J\'unior, M. Assafin, L. Beauvalet et al.]{A. R. Gomes-J\'unior$^{1}$\thanks{E-mail: altair08@astro.ufrj.br},
M. Assafin$^{1,\dag}$,
L. Beauvalet$^{2,3}$,
J. Desmars$^{4}$,\newauthor
R. Vieira-Martins$^{1,2,5,\dag}$,
J. I. B. Camargo$^{2,5}$,
B. E. Morgado$^{1,2}$
F. Braga-Ribas$^{2,6}$
\\
$^{1}$Observat\'orio do Valongo/UFRJ, Ladeira Pedro Ant\^onio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
$^{2}$Observat\'orio Nacional/MCTI, R. General Jos\'e Cristino 77, CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
$^{3}$Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau 75014 Paris, France\\
$^{4}$Institut de m\'ecanique c\'eleste et de calcul des \'eph\'em\'erides - Observatoire de Paris, UMR 8028 du CNRS, 77 Av. Denfert-Rochereau,\\ 75014 Paris, France\\
$^{5}$Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ 20921-400, Brazil\\
$^{6}$Federal University of Technology - Paran\'a (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil
}
\begin{document}
\newcommand{\noccs}{5442 } % total number of occultations
\defcitealias{GomesJunior2015}{GJ15}

@@

\begin{abstract}
%Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, shape, albedo and composition, to trace back their true origin. The best ground-based technique to determine size and shape, and thus constrain the albedo and in a broader sense composition, is the observation of stellar occultations by these objects but this demands a very precise ephemeris based on new and good positions.
%\cite{GomesJunior2015} published 6523 observations obtained between 1992 and 2014 for 18 irregular satellites with an estimated error in the positions of about 60 to 80 mas.
%Based on this observations, we aimed to generate new numerical integration of the orbits of eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda. This is a short-time ephemeris that corrected systematic errors from previous ephemeris (JPL JUP300). For Phoebe we updated ephemeris from \cite{Desmars2013} using 75\% more positions than the previous one.
%Also Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018. This creates a good opportunity to predict stellar occultations due to the large number of events that can be predicted in these periods. Thus, we aimed to predict stellar occultations for the same 9 satellites cited above. Using the UCAC4 catalogue we managed to identify \noccs candidates of stellar occultations between the period of January, 2016 and December, 2020.
%We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.
\cite{GomesJunior2015} published 6523 new positions for 18 irregular satellites from observations made between 1992 and 2014 with an estimated error of about 60 to 80 mas. Based on this set of original positons, we performed numerical integrations for deriving new orbits for the eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda. This corrected systematic errors and resulted in an ephemeris more accurate/precise than JPL JUP300 for these bodies at least in the immediate future. For Phoebe we updated the ephemeris from \cite{Desmars2013} using 75\% more positions than the previous one. Due to their orbital configurations, it is common belief that the irregular satellites were captured by the giant planets in the early Solar System, but there is no consensus for a single model explaining where they were formed. Size, shape, albedo and composition would help a lot to trace back their true origin, but these physical parameters are yet poorly known for irregular satellites. The observation of stellar occultations would allow for the determination of such parameters. Indeed Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018, improving a lot the chances of observing such events in the near future. Using the derived ephemerides and the UCAC4 catalogue we managed to identify \noccs candidate stellar occultations between January 2016 and December 2020 for the 9 irregular satellites studied here. We discussed how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.
\end{abstract}

\begin{keywords}
occultations - planets and satellites: general - planets and satellites: individual: Jovian and Saturnian irregular satellites - planets and satellites: individual: Triton
\end{keywords}

@@

\cite{Grav2003} and \cite{Grav2007} showed that the irregular satellites from the giant planets have their colors and spectral slopes similar to C-, D- and P-type asteroids, Centaurs and trans-neptunian objects (TNOs). This suggests that they may have come from different locations in the early solar system.

@@

We present in this paper new numerical integration of the orbits of the 8 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke, Sinope and Leda) using only the positions obtained by \citetalias{GomesJunior2015}. We aim to eliminate the systematic errors in the JPL ephemeris (see Sec. \ref{Sec: integration}). %For an object at the distance of Jupiter, this represents an error langer than 700 km in the shadow path of an occultation.
For Phoebe, we updated the ephemeris of \cite{Desmars2013} using the observations of \citetalias{GomesJunior2015}, \cite{Peng2015}, observations from Minor Planet Center and observations from Flagstaff.

@@

\begin{table}
\caption{\label{Tab: satellite-diameter} Estimated diameter of the satellites and correspondent apparent diameter}
\begin{center}
\begin{tabular}{lccc}
\hline  \hline
\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & mas$^ {a}$  & km & Ref. \tabularnewline
\hline
Ananke & 8 & 29 & 1 \tabularnewline
Carme & 13 & 46 & 1 \tabularnewline
Elara & 24 & 86 & 1 \tabularnewline
Himalia & 41 & $(150\times120) \pm 20^{b}$ & 2 \tabularnewline
Leda & 5 & 20 & 1 \tabularnewline
Lysithea & 10 & 36 & 1 \tabularnewline
Pasiphae & 17 & 62 & 1 \tabularnewline
Sinope & 10 & 37 & 1 \tabularnewline
\hdashline
Phoebe & 32 & $212 \pm 1.4^{b}$ & 3 \tabularnewline
%\hdashline
%Nereid & 15 & $340 \pm 50^{c}$ & 4 \tabularnewline
%Triton & 124 & $2707 \pm 2.0^{c}$ & 5 \tabularnewline
\hline
\end{tabular}
\end{center}
References: (1) \cite{Rettig2001}; (2) \cite{Porco2003}; (3) \cite{Thomas2010}..\\
$^{a}${Using a mean distance from Jupiter of 5 AU, from Saturn of 9 AU and from Neptune of 30 AU.}
$^{b}${From Cassini observations.}
\par
\end{table}

@@

In section \ref{Sec: integration} the new determination of the orbits. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites, including some tests made to check the accuracy of the predictions. The final discussion is presented in section \ref{Sec: discussion}.

@@

Here we compute new orbits based on the observations published in \citetalias{GomesJunior2015}. First, because the reduction was made with a consistent and precise stellar catalogue and with a robust astrometry \citep[PRAIA,][]{Assafin2011}. Second, besides recent observations, this consistent set of numerous and precise positions covers many orbital periods at many distinct orbital plane sights, so that the orbital inclinations along with all other orbital elements could be satisfactorily derived without the need of further position sets. For these reasons, only this set of positions was used for the satellites of Jupiter.

Due to the context of this work regarding to stellar occultations, the orbit fitting procedures used aimed primarily to derive precise ephemerides for the near future. The procedures allow for the improvement of orbits as new observations are added.

\subsection{Special tailored Ephemerides (STE) for Jupiter irregular satellites}

The last observations used to develop JPL current ephemeris of the irregular satellites of Jupiter were obtained in 2012 \citep{Jacobson2012}. As a result, the errors in the JPL ephemeris for the current epoch are large enough to prevent accurate predictions of stellar occultations without any corrections.

@@

The obtained ephemeris is hereafter refered as STE, for special-tailored ephemeris. The initial osculating elements at the origin of integration are presented in Table \ref{Tab: sat_ell}.

\begin{table*}
\caption{Initial osculating elements for Jupiter irregular satellites at JD 2451545.0.
 }\label{Tab: sat_ell}
\begin{center}
\begin{tabular}{ccccccc}
\hline\hline
Satellite & a (km) & e & I$\degr$ & $\Omega\degr$ & $\omega\degr$ & $v\degr$ \\ 
\hline
Himalia &   11372100 $\pm$ 500    &    0.166 $\pm$ 0.002      &   45.14 $\pm$ 0.15      &   39.77 $\pm$ 0.19      &   351.48 $\pm$ 0.46      &   97.35 $\pm$ 0.48    \\
Elara &   11741170 $\pm$ 690  &      0.222 $\pm$ 0.002      &   28.64 $\pm$ 0.18      &   68.42 $\pm$ 0.43      &   179.82 $\pm$ 0.56      &   339.08 $\pm$ 0.82  \\
Lysithea &   11739900 $\pm$ 1300  &      0.136 $\pm$ 0.004      &    51.12 $\pm$ 0.27     &   5.53 $\pm$ 0.52      &   53.0 $\pm$ 1.5      &   318.9 $\pm$ 2.0   \\
Leda &   11140300  $\pm$ 4300  &     0.173  $\pm$ 0.007     &   16.15  $\pm$ 0.75    &   272.6  $\pm$ 1.7    &   212.2  $\pm$ 3.6          &   218.8  $\pm$ 3.2  \\
Pasiphae &  23425000  $\pm$ 5000    &     0.379  $\pm$ 0.001       &   152.44 $\pm$ 0.10      &   284.59 $\pm$ 0.21      &   135.96 $\pm$ 0.19      &   236.97 $\pm$ 0.16 \\
Sinope &   22968800 $\pm$ 5200   &     0.316 $\pm$ 0.002      &   157.76 $\pm$ 0.12      &   256.62 $\pm$ 0.55      &   298.38 $\pm$ 0.55      &   167.57 $\pm$ 0.19    \\
Carme &   24202924 $\pm$ 4800      &  0.242 $\pm$ 0.001      &   147.13 $\pm$ 0.10      &   154.01 $\pm$ 0.25      &   47.90 $\pm$ 0.29      &   234.41 $\pm$ 0.19  \\
Ananke &  21683800  $\pm$ 7200  &     0.380 $\pm$ 0.002      &   172.29 $\pm$ 0.20      &   56.9 $\pm$ 1.2      &   123.3 $\pm$ 1.2      &   231.24 $\pm$ 0.21  \\
%Themisto  & 7393800    &     0.198       &   25.77  &   220.0     &   216.3    &   262.1     \\
\hline
\end{tabular} 
\end{center}
\begin{flushleft}
\textbf{Notes}: a: semimajor axis; e : excentricity; I: inclination relative to the equatorial reference plane J2000; $\Omega$: longitude of the ascending node; $\omega$: argument of periapsis; $v$: true anomaly.
\end{flushleft}
\end{table*}

All the orbits determined for the satellites show satisfying residuals. The residuals are smaller than those obtained with JPL ephemeris, which was expected because the accuracy of an ephemeris decreases when we get further from the time of observations. The main risk of divergence over time comes from the possible absence of long-term effects when fitting to a short timespan of observations. If that were the case, our ephemeris would diverge too quickly to be of any use. JPL ephemerides are fitted over all the available observations. As a result, they will diverge less quickly than our own. Though they are no longer precise enough for our use, they remain a precious reference to identify whether our own model presents a quick divergence.

@@

Fig. \ref{Fig: JPL-STE} displays the offsets of the positions published by \citetalias{GomesJunior2015} for the satellite Carme in declination relative to the STE ephemeris, to \cite{Jacobson2012} JUP300 JPL ephemeris and \cite{Emelyanov2008} ephemeris.We see that the systematic JPL ephemeris offsets pointed out by \citetalias{GomesJunior2015} are reduced with our ephemeris, as expected.

\begin{figure}
\includegraphics[width=8.8cm]{figures/Carme_ephemeris.eps}
\caption{Offsets in declination of the positions published by \protect\citetalias{GomesJunior2015} for Carme. The red "x" relative to the special-tailored ephemeris, the blue "+" relative to the JUP300 JPL ephemeris and the green dot relative to \protect\cite{Emelyanov2008}. As expected, the ephemeris systematic errors pointed out by \protect\citetalias{GomesJunior2015} are reduced with the STE ephemeris. \label{Fig: JPL-STE}}
\end{figure}

In Table \ref{Tab:STE-error} it is presented the mean offsets and the respective standard deviation of the \citetalias{GomesJunior2015} positions relative to the same 3 ephemeris as above. We can see that the mean offsets are greatly reduced with the STE ephemeris.

\begin{table*}
\caption{\label{Tab:STE-error} Mean offsets and standard deviation of the \protect\citetalias{GomesJunior2015} positions relative to the STE, \protect\cite{Jacobson2012} and \protect\cite{Emelyanov2008} ephemeris .}
\begin{centering}
\begin{tabular}{l|cc|cc|cc}
\hline  \hline
%\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
 & \multicolumn{2}{|c|}{STE} & \multicolumn{2}{|c|}{JPL} & \multicolumn{2}{|c}{Eme2008}\tabularnewline
Satellite  & RA & DEC & RA & DEC & RA & DEC \tabularnewline
\hline
Ananke & $-10\pm90$ & $3\pm73$ & $60\pm127$ & $-108\pm99$ & $101\pm180$ & $-107\pm120$ \tabularnewline
Carme & $14\pm73$ & $-1\pm51$ & $-3\pm114$ & $-80\pm102$ & $-6\pm108$ & $-45\pm80$ \tabularnewline
Elara & $3\pm92$ & $-12\pm57$ & $20\pm92$ & $-50\pm69$ & $23\pm94$ & $-83\pm81$ \tabularnewline
Himalia & $-15\pm66$ & $-7\pm54$  & $-19\pm80$  & $-11\pm52$ & $-18\pm72$  & $-13\pm53$ \tabularnewline
Leda & $-9\pm67$ & $-8\pm77$ & $60\pm117$ & $-13\pm95$ & $166\pm162$ & $92\pm95$ \tabularnewline
Lysithea & $15\pm79$ & $-21\pm68$ & $40\pm92$ & $-43\pm77$ & $117\pm193$ & $-76\pm185$ \tabularnewline
Pasiphae & $4\pm89$ & $-16\pm57$ & $-17\pm130$ & $-82\pm85$ & $-10\pm102$ & $-54\pm74$ \tabularnewline
Sinope & $9\pm79$ & $-4\pm47$ & $10\pm228$ & $-35\pm76$ & $11\pm227$ & $-52\pm63$ \tabularnewline
\hline
\end{tabular}
\par \end{centering}
\end{table*}

@@

For the specific case of Phoebe, the ninth satellite of Saturn, we have updated the ephemeris published in \cite{Desmars2013}. The new ephemeris (PH15) used the same dynamical model, including the perturbations of the Sun and the eight planets, the eight major satellites of Saturn and the $J_2$ parameter. The observations used to fit the model are identical to \cite{Desmars2013} (including 223 Cassini observations) with additional observations from \citetalias{GomesJunior2015}, \cite{Peng2015}, observations from Minor Planet Circulars between 2012 and 2014 (available on the Natural Satellite Data Center\footnote{\url{http://lnfm1.sai.msu.ru/neb/nss/bsapoouf.htm}}), and observations from Flagstaff \citep{NOFS} between 2012 and 2014. It represents a total number of 5886 observations from 1898 to 2014. In contrast, in \cite{Desmars2013} was used 3367 observations from 1898 to 2012. It represents an increase of almost 75\% in the number of observations.

In Fig. \ref{Fig:eph-Phoebe} we compare our ephemeris (PH15) with the SAT375 JPL\footnote{Jacobson, R.A. 2015-Feb-27. "Satellite Ephemeris: SAT375", JPL Satellite Ephemeris File Release, \url{ftp://ssd.jpl.nasa.gov/pub/eph/satellites/nio/LINUX_PC/sat375l.txt}} ephemeris. The difference between them is smaller than 30 mas (< 10 mas in Declination). This difference is smaller than the apparent diameter of Phoebe (see Table \ref{Tab: satellite-diameter})

\begin{figure}
\begin{centering}
\includegraphics[width=8.8cm]{figures/Phoebe.eps} 
\caption{Comparison between the PH15 and SAT375 JPL ephemeris for the satellite Phoebe.}
\label{Fig:eph-Phoebe}
\end{centering}
\end{figure}

We computed the precision of the PH15 ephemeris considering three sets of observations: all the positions available; only the positions of \citetalias{GomesJunior2015}; and all the positions without \citetalias{GomesJunior2015}. In Fig. \ref{Fig:err-Phoebe} we show the comparison between them for the time spam 2016-2021 in angular separation of Saturn. It is possible to see that even considering only the positions of \citetalias{GomesJunior2015}, the estimated error of the ephemeris is smaller than 12 mas. The computed precision does not take into account the precision in the position of Saturn.

\begin{figure}
\begin{centering}
\includegraphics[width=8.8cm]{figures/PH15_err_angsep.eps}  
\caption{Comparison of the precision in angular separation of Saturn for the PH15 ephemeris where three different sets of positions were used to compute the ephemeris: all observations; only positions of \citetalias{GomesJunior2015}; and all the positions without \citetalias{GomesJunior2015}.}
\label{Fig:err-Phoebe}
\end{centering}
\end{figure}

>>>>

\title[New orbits and predictions of stellar occultations by irregular satellites up to 2020]{New orbits of irregular satellites based on thousands of new observations and predictions for stellar occultations up to 2020}
\author[A. R. Gomes-J\'unior, M. Assafin, L. Beauvalet et al.]{A. R. Gomes-J\'unior$^{1}$\thanks{E-mail: altair08@astro.ufrj.br},
M. Assafin$^{1,\dag}$,
L. Beauvalet$^{2,3}$,
J. Desmars$^{4}$,\newauthor
R. Vieira-Martins$^{1,2,5,\dag}$,
J. I. B. Camargo$^{2,5}$,
B. E. Morgado$^{1,2}$
F. Braga-Ribas$^{2,6}$
\\
$^{1}$Observat\'orio do Valongo/UFRJ, Ladeira Pedro Ant\^onio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
$^{2}$Observat\'orio Nacional/MCTI, R. General Jos\'e Cristino 77, CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
$^{3}$Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau 75014 Paris, France\\
$^{4}$Institut de m\'ecanique c\'eleste et de calcul des \'eph\'em\'erides - Observatoire de Paris, UMR 8028 du CNRS, 77 Av. Denfert-Rochereau,\\ 75014 Paris, France\\
$^{5}$Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ 20921-400, Brazil\\
$^{6}$Federal University of Technology - Paran\'a (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil
}
\begin{document}
\newcommand{\noccs}{5442 } % total number of occultations
\defcitealias{GomesJunior2015}{GJ15}

@@

\begin{abstract}
%Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, shape, albedo and composition, to trace back their true origin. The best ground-based technique to determine size and shape, and thus constrain the albedo and in a broader sense composition, is the observation of stellar occultations by these objects but this demands a very precise ephemeris based on new and good positions.
%\cite{GomesJunior2015} published 6523 observations obtained between 1992 and 2014 for 18 irregular satellites with an estimated error in the positions of about 60 to 80 mas.
%Based on this observations, we aimed to generate new numerical integration of the orbits of eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda. This is a short-time ephemeris that corrected systematic errors from previous ephemeris (JPL JUP300). For Phoebe we updated ephemeris from \cite{Desmars2013} using 75\% more positions than the previous one.
%Also Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018. This creates a good opportunity to predict stellar occultations due to the large number of events that can be predicted in these periods. Thus, we aimed to predict stellar occultations for the same 9 satellites cited above. Using the UCAC4 catalogue we managed to identify \noccs candidates of stellar occultations between the period of January, 2016 and December, 2020.
%We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.
\cite{GomesJunior2015} published 6523 new positions for 18 irregular satellites from observations made between 1992 and 2014 with an estimated error of about 60 to 80 mas. Based on this set of original positions, we performed numerical integrations for deriving new orbits for the eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda. This corrected systematic errors and resulted in an ephemeris more accurate/precise than JPL jup300 for these bodies at least in the immediate future. For Phoebe we updated the ephemeris from \cite{Desmars2013} using 75\% more positions than the previous one. Due to their orbital characteristics, it is common belief that the irregular satellites were captured by the giant planets in the early Solar System, but there is no consensus for a single model explaining where they were formed. Size, shape, albedo and composition would help a lot to trace back their true origin, but these physical parameters are yet poorly known for irregular satellites. The observation of stellar occultations would allow for the determination of such parameters. Indeed Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018, improving a lot the chances of observing such events in the near future. Using the derived ephemerides and the UCAC4 catalogue we managed to identify \noccs candidate stellar occultations between January 2016 and December 2020 for the 9 irregular satellites studied here. We discussed how the successful observation of a stellar occultation by these objects is possible and present some of those potential occultations.
\end{abstract}

\begin{keywords}
occultations - planets and satellites: general - planets and satellites: individual: Jovian and Saturnian irregular satellites
\end{keywords}

@@

\cite{Grav2003} and \cite{Grav2007} showed that the irregular satellites from the giant planets have their colors and spectral slopes similar to C-, D- and P-type asteroids, Centaurs and trans-neptunian objects (TNOs). This suggests that they may have come from different locations in the early Solar System.

@@

We present in this paper new numerical integration of the orbits of the 8 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke, Sinope and Leda) using only the positions obtained by \citetalias{GomesJunior2015}%. We aim to eliminate the systematic errors in the JPL ephemeris
 (see Sec. \ref{Sec: integration}). %For an object at the distance of Jupiter, this represents an error langer than 700 km in the shadow path of an occultation.
For Phoebe, we updated the ephemeris of \cite{Desmars2013} using the observations of \citetalias{GomesJunior2015}, \cite{Peng2015}, observations from Minor Planet Center and observations from Flagstaff.

@@

\begin{table}
\caption{\label{Tab: satellite-diameter} Estimated diameter of the satellites and correspondent apparent diameter}
\begin{center}
\begin{tabular}{lccc}
\hline  \hline
\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & mas$^ {a}$  & km & Ref. \tabularnewline
\hline
Ananke & 8 & 29 & 1 \tabularnewline
Carme & 13 & 46 & 1 \tabularnewline
Elara & 24 & 86 & 1 \tabularnewline
Himalia & 41 & $(150\times120) \pm 20^{b}$ & 2 \tabularnewline
Leda & 5 & 20 & 1 \tabularnewline
Lysithea & 10 & 36 & 1 \tabularnewline
Pasiphae & 17 & 62 & 1 \tabularnewline
Sinope & 10 & 37 & 1 \tabularnewline
\hdashline
Phoebe & 32 & $212 \pm 1.4^{b}$ & 3 \tabularnewline
%\hdashline
%Nereid & 15 & $340 \pm 50^{c}$ & 4 \tabularnewline
%Triton & 124 & $2707 \pm 2.0^{c}$ & 5 \tabularnewline
\hline
\end{tabular}
\end{center}
References: (1) \cite{Rettig2001}; (2) \cite{Porco2003}; (3) \cite{Thomas2010}.\\
$^{a}${Using a mean distance from Jupiter of 5 AU, from Saturn of 9 AU and from Neptune of 30 AU.}\\
$^{b}${From Cassini observations.}
\par
\end{table}

@@

In section \ref{Sec: integration}, we present the new determination of the orbits. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites, including some tests made to check the accuracy of the predictions. The final discussion is presented in section \ref{Sec: discussion}.

@@

Here we compute new orbits based on the observations published in \citetalias{GomesJunior2015}. First, because the reduction was made with a consistent and precise stellar catalogue and with a robust astrometry \citep[PRAIA,][]{Assafin2011}. Second, besides recent observations, this consistent set of numerous and precise positions covers many orbital periods at many distinct orbital plane sights, allowing to fully constrain the orbit for the short time span explored in this work. For these reasons, only this set of positions was used for the satellites of Jupiter.

Due to the context of this work regarding to stellar occultations, the orbit fitting procedures used aimed primarily to derive precise ephemerides for the near future. The procedures allow for the improvement of orbits as recent observations are added.

\subsection{Special-Tailored Ephemerides (STE) for Jupiter irregular satellites}

The last observations used to develop current JPL ephemeris of the irregular satellites of Jupiter were obtained in 2012 \citep{Jacobson2012}. As a result, the errors in the JPL ephemeris for the current epoch may be probably too large to prevent accurate predictions of stellar occultations without any corrections.

@@

The obtained ephemeris is hereafter refered as STE, for Special-Tailored Ephemeris. The initial osculating elements at the origin of integration are presented in Table \ref{Tab: sat_ell}.

\begin{table*}
\caption{Initial osculating elements for Jupiter irregular satellites at JD 2451545.0 with respect to the center of Jupiter.
 }\label{Tab: sat_ell}
\begin{center}
\begin{tabular}{ccccccc}
\hline\hline
Satellite & a (km) & e & I$\degr$ & $\Omega\degr$ & $\omega\degr$ & $v\degr$ \\ 
\hline
Himalia &   11372100 $\pm$ 500    &    0.166 $\pm$ 0.002      &   45.14 $\pm$ 0.15      &   39.77 $\pm$ 0.19      &   351.48 $\pm$ 0.46      &   97.35 $\pm$ 0.48    \\
Elara &   11741170 $\pm$ 690  &      0.222 $\pm$ 0.002      &   28.64 $\pm$ 0.18      &   68.42 $\pm$ 0.43      &   179.82 $\pm$ 0.56      &   339.08 $\pm$ 0.82  \\
Lysithea &   11739900 $\pm$ 1300  &      0.136 $\pm$ 0.004      &    51.12 $\pm$ 0.27     &   5.53 $\pm$ 0.52      &   53.0 $\pm$ 1.5      &   318.9 $\pm$ 2.0   \\
Leda &   11140300  $\pm$ 4300  &     0.173  $\pm$ 0.007     &   16.15  $\pm$ 0.75    &   272.6  $\pm$ 1.7    &   212.2  $\pm$ 3.6          &   218.8  $\pm$ 3.2  \\
Pasiphae &  23425000  $\pm$ 5000    &     0.379  $\pm$ 0.001       &   152.44 $\pm$ 0.10      &   284.59 $\pm$ 0.21      &   135.96 $\pm$ 0.19      &   236.97 $\pm$ 0.16 \\
Sinope &   22968800 $\pm$ 5200   &     0.316 $\pm$ 0.002      &   157.76 $\pm$ 0.12      &   256.62 $\pm$ 0.55      &   298.38 $\pm$ 0.55      &   167.57 $\pm$ 0.19    \\
Carme &   24202924 $\pm$ 4800      &  0.242 $\pm$ 0.001      &   147.13 $\pm$ 0.10      &   154.01 $\pm$ 0.25      &   47.90 $\pm$ 0.29      &   234.41 $\pm$ 0.19  \\
Ananke &  21683800  $\pm$ 7200  &     0.380 $\pm$ 0.002      &   172.29 $\pm$ 0.20      &   56.9 $\pm$ 1.2      &   123.3 $\pm$ 1.2      &   231.24 $\pm$ 0.21  \\
%Themisto  & 7393800    &     0.198       &   25.77  &   220.0     &   216.3    &   262.1     \\
\hline
\end{tabular} 
\end{center}
\begin{flushleft}
\textbf{Notes}: a: semimajor axis; e : excentricity; I: inclination relative to the equatorial reference plane J2000; $\Omega$: longitude of the ascending node; $\omega$: argument of periapsis; $v$: true anomaly.
\end{flushleft}
\end{table*}

All the orbits determined for the satellites show satisfying residuals. The residuals are smaller than those obtained with JPL ephemeris, which was expected because the accuracy of an ephemeris decreases when we get further from the time of observations. The main risk of divergence over time comes from the possible absence of long-term effects when fitting to a short timespan of observations. If that was the case, our ephemeris would diverge too quickly to be of any use. JPL ephemerides are fitted over all the available observations. As a result, they will diverge less quickly than our own. Though they are not as much precise as required for the use, they remain a precious reference to identify whether our own model presents a quick divergence.

@@

Fig. \ref{Fig: JPL-STE} displays the offsets of the positions published by \citetalias{GomesJunior2015} for the satellite Carme in declination relative to the STE ephemeris, to \cite{Jacobson2012} jup300 JPL ephemeris and \cite{Emelyanov2008} ephemeris.We see that the systematic JPL ephemeris offsets pointed out by \citetalias{GomesJunior2015} are reduced with our ephemeris, as expected.

\begin{figure}
\includegraphics[width=8.8cm]{figures/Carme_ephemeris.eps}
\caption{Offsets in declination of the positions published by \protect\citetalias{GomesJunior2015} for Carme. The red "x" relative to the special-tailored ephemeris, the blue "+" relative to the jup300 JPL ephemeris and the green dot relative to \protect\cite{Emelyanov2008}. As expected, the ephemeris systematic errors pointed out by \protect\citetalias{GomesJunior2015} are reduced with the STE ephemeris. \label{Fig: JPL-STE}}
\end{figure}

In Table \ref{Tab:STE-error} it is presented the mean offsets and the respective standard deviation of the \citetalias{GomesJunior2015} positions relative to the same 3 ephemeris as above. We can see that the mean offsets as well as most of their standard deviations are greatly reduced with the STE ephemeris.

\begin{table*}
\caption{\label{Tab:STE-error} Mean offsets and standard deviation of the \protect\citetalias{GomesJunior2015} positions relative to the STE, \protect\cite{Jacobson2012} and \protect\cite{Emelyanov2008} ephemeris .}
\begin{centering}
\begin{tabular}{l|cc|cc|cc}
\hline  \hline
%\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
 & \multicolumn{2}{|c|}{STE} & \multicolumn{2}{|c|}{JPL} & \multicolumn{2}{|c}{Eme2008}\tabularnewline
Satellite  & $\Delta\alpha\cos\delta$ (mas) & $\Delta\delta$ (mas) & $\Delta\alpha\cos\delta$ (mas) & $\Delta\delta$ (mas) & $\Delta\alpha\cos\delta$ (mas) & $\Delta\delta$ (mas) \tabularnewline
\hline
Ananke & $-10\pm90$ & $3\pm73$ & $60\pm127$ & $-108\pm99$ & $101\pm180$ & $-107\pm120$ \tabularnewline
Carme & $14\pm73$ & $-1\pm51$ & $-3\pm114$ & $-80\pm102$ & $-6\pm108$ & $-45\pm80$ \tabularnewline
Elara & $3\pm92$ & $-12\pm57$ & $20\pm92$ & $-50\pm69$ & $23\pm94$ & $-83\pm81$ \tabularnewline
Himalia & $-15\pm66$ & $-7\pm54$  & $-19\pm80$  & $-11\pm52$ & $-18\pm72$  & $-13\pm53$ \tabularnewline
Leda & $-9\pm67$ & $-8\pm77$ & $60\pm117$ & $-13\pm95$ & $166\pm162$ & $92\pm95$ \tabularnewline
Lysithea & $15\pm79$ & $-21\pm68$ & $40\pm92$ & $-43\pm77$ & $117\pm193$ & $-76\pm185$ \tabularnewline
Pasiphae & $4\pm89$ & $-16\pm57$ & $-17\pm130$ & $-82\pm85$ & $-10\pm102$ & $-54\pm74$ \tabularnewline
Sinope & $9\pm79$ & $-4\pm47$ & $10\pm228$ & $-35\pm76$ & $11\pm227$ & $-52\pm63$ \tabularnewline
\hline
\end{tabular}
\par \end{centering}
\end{table*}

@@

For the specific case of Phoebe, the ninth satellite of Saturn, we have updated the ephemeris published in \cite{Desmars2013}. The new ephemeris (PH15) used the same dynamical model, including the perturbations of the Sun and the eight planets, the eight major satellites of Saturn and the $J_2$ parameter. The observations used to fit the model are identical to \cite{Desmars2013} (including 223 Cassini observations) with additional observations from \citetalias{GomesJunior2015}, \cite{Peng2015}, observations from Minor Planet Circulars between 2012 and 2014 (available on the Natural Satellite Data Center\footnote{\url{http://lnfm1.sai.msu.ru/neb/nss/bsapoouf.htm}}), and observations from Flagstaff \citep{NOFS} between 2012 and 2014. It represents a total number of 5886 observations from 1898 to 2014. In contrast, in \cite{Desmars2013} was used 3367 observations from 1898 to 2012. This represents an increase of almost 75\% in the number of observations, mainly with recent observations which is required for our purpose.

In Fig. \ref{Fig:eph-Phoebe} we compare our ephemeris (PH15) with the sat375 JPL\footnote{Jacobson, R.A. 2015-Feb-27. "Satellite Ephemeris: sat375", JPL Satellite Ephemeris File Release, \url{ftp://ssd.jpl.nasa.gov/pub/eph/satellites/nio/LINUX_PC/sat375l.txt}} ephemeris. The difference between them is smaller than 30 mas (< 10 mas in Declination). This difference is smaller than the apparent diameter of Phoebe (see Table \ref{Tab: satellite-diameter})

\begin{figure}
\begin{centering}
\includegraphics[width=8.8cm]{figures/Phoebe.eps} 
\caption{Comparison between the PH15 and sat375 JPL ephemeris for the satellite Phoebe.}
\label{Fig:eph-Phoebe}
\end{centering}
\end{figure}

We computed the precision of the PH15 ephemeris considering three sets of observations: all the positions available; only the positions of \citetalias{GomesJunior2015}; and all the positions without \citetalias{GomesJunior2015}. In Fig. \ref{Fig:err-Phoebe} we show the comparison between them for the time span 2016-2021 in angular separation\footnote{\label{ft-angsep}The precision in angular separation is computed  by $\sigma_s = \sqrt{\sigma^2_{\alpha} \cos^2 \delta + \sigma^2_{\delta}}$ where $\sigma_{\alpha}$ and $\sigma_{\delta}$ are the precision in right ascension and declination.} of Saturn. It is possible to see that even considering only the positions of \citetalias{GomesJunior2015}, the estimated error of the ephemeris is smaller than 12 mas. The computed precision does not take into account the precision in the position of Saturn.
